### 2013-11-06 src2/thesis.tex
update

@@

\begin{small}
\begin{Verbatim}[samepage=true]
int main(void) {

@@

>>>>

Each MP of the GPU can manage a large number of threads; on today's GPUs up to 
2048 threads can run on a single MP. The GPU schedules threads in groups of 32, called 
{\em Warps}, that are executed in lock-step (SIMD style execution). Threads are 
also divided into {\em Blocks}; the threads within a block can communicate using the 
shared memory. The maximum block size is currently 1024 threads. Since 
the number of threads in a block can be higher than the warp size, a barrier 
synchronisation mechanism exists to ensure that threads within a block 
can communicate safely via the shared memory. Blocks are also grouped into a grid that 
is the collection of blocks executing the same program.

\subsection{GPU programming}

The CUDA programming model reflects the hierarchical architecture of the GPU. The 
programmer writes what are called \emph{kernels}; in CUDA jargon a kernel refers to a sequential 
C program that is parameterised on a thread's identity. The kernel program is executed 
in parallel across all the threads of a block and many such blocks of threads can execute 
in parallel on the available MPs. Kernels may use shared memory that is local to a block; 
the programmer needs to decompose the algorithm and data in such a way that communication 
will be local to the blocks.

The program below is a very simple example of a CUDA kernel that computes elementwise sums.

\noindent The kernel takes two input arrays, {\tt i1} and {\tt i2}; the result is stored 
in {\tt r}. Note how the kernel makes use of three variables, {\tt threadIdx.x}, 
{\tt blockIdx.x} and {\tt gridDim.x}. These variables identify a thread uniquely within 
a grid. Here they are used to obtain an index into the provided arrays. The dimensions of 
the blocks and of the grid is set by a controling program upon launch of the kernel. This 
controling program is run by the CPU of the computer housing the GPU. the kernel is launched 
using the following syntax:

\begin{small} 
\begin{Verbatim}[samepage=true]
kernel<<<1,32,0>>>(d_i1,d_i2,d_r);
\end{Verbatim}
\end{small}

\noindent The line of code above launches the kernel on one block of 32 threads. The zero 
indicates that the kernel uses no shared memory. If the kernel did use shared memory, the 
amount needed should have been specified there. Figure~\ref{fig:CUDACODE1} shows the complete 
sequence of commands for a minimal example using the kernel specified above.

In Obsidian the same kernel (the local computation) is implemented as follows:

\begin{small}
\begin{Verbatim}[samepage=true]

\end{Verbatim]
\end{small}

@@

\begin{figure} 
\begin{small}
\begin{Verbatim}[samepage=true]
int main(void) {

@@

\caption{Host code for launching a simple CUDA kernel on the GPU.}
\label{fig:CUDACODE1}

\end{figure}
